\chapter{How Robin Hood Met Little John}

\begin{quote}
``O here is my hand,'' the stranger reply'd,\\
``I'll serve you with all my whole heart.\\
My name is John Little, a man of good mettle,\\
Ne'er doubt me for I'll play my part.''

``His name shall be altered,'' quoth William Stutely,\\
``And I will his godfather be:\\
Prepare then a feast, and none of the least,\\
For we will be merry,'' quoth he.
\end{quote}

\lettrine{A}{ll} that summer Robin Hood and his merry men roamed in
Sherwood Forest, and the fame of their deeds ran abroad in the land. The
Sheriff of Nottingham waxed wroth at the report, but all his traps and
excursions failed to catch the outlaws. The poor people began by fearing
them, but when they found that the men in Lincoln green who answered
Robin Hood's horn meant them no harm, but despoiled the oppressor to
relieve the oppressed, they `gan to have great liking for them. And the
band increased by other stout hearts till by the end of the summer
fourscore good men and true had sworn fealty.

But the days of quiet which came on grew irksome to Robin's adventurous
spirit. Up rose he, one gay morn, and slung his quiver over his
shoulders.

``This fresh breeze stirs the blood, my lads,'' quoth he, ``and I would
be seeing what the gay world looks like in the direction of Nottingham
town. But tarry ye behind in the borders of the forest, within earshot
of my bugle call.''

Thus saying he strode merrily forward to the edge of the wood, and
paused there a moment, his agile form erect, his brown locks flowing and
his brown eyes watching the road; and a goodly sight he made, as the
wind blew the ruddy color into his cheeks.

The highway led clear in the direction of the town, and thither he
boldly directed his steps. But at a bend in the road he knew of a
by-path leading across a brook which made the way nearer and less open,
into which he turned. As he approached the stream he saw that it had
become swollen by recent rains into quite a pretty torrent. The log
foot-bridge was still there, but at this end of it a puddle intervened
which could be crossed only with a leap, if you would not get your feet
wet.

But Robin cared little for such a handicap. Taking a running start, his
nimble legs carried him easily over and balanced neatly upon the end of
the broad log. But he was no sooner started across than he saw a tall
stranger coming from the other side. Thereupon Robin quickened his pace,
and the stranger did likewise, each thinking to cross first. Midway they
met, and neither would yield an inch.

``Give way, fellow!'' roared Robin, whose leadership of a band, I am
afraid, had not tended to mend his manners.

The stranger smiled. He was almost a head taller than the other.

``Nay,'' he retorted, ``fair and softly! I give way only to a better man
than myself.''

``Give way, I say'', repeated Robin, ``or I shall have to show you a
better man.''

His opponent budged not an inch, but laughed loudly. ``Now by my
halidom!'' he said good-naturedly, ``I would not move after hearing that
speech, even if minded to it before; for this better man I have sought
my life long. Therefore show him to me, an it please you.''

``That will I right soon,'' quoth Robin. ``Stay you here a little while,
till I cut me a cudgel like unto that you have been twiddling in your
fingers.'' So saying he sought his own bank again with a leap, laid
aside his long bow and arrows, and cut him a stout staff of oak,
straight, knotless, and a good six feet in length. But still it was a
full foot shorter than his opponent's. Then back came he boldly.

``I mind not telling you, fellow,'' said he, ``that a bout with archery
would have been an easier way with me. But there are other tunes in
England besides that the arrow sings.'' Here he whirred the staff about
his head by way of practice. ``So make you ready for the tune I am about
to play upon your ribs. Have at you! One, two--''

``Three!'' roared the giant smiting at him instantly.

Well was it for Robin that he was quick and nimble of foot; for the blow
that grazed a hair's breadth from his shoulder would have felled an ox.
Nevertheless while swerving to avoid this stroke, Robin was poising for
his own, and back came he forthwith--whack!

Whack! parried the other.

Whack! whack! whack! whack!

The fight waxed fast and furious. It was strength pitted against
subtlety, and the match was a merry one. The mighty blows of the
stranger went whistling around Robin's ducking head, while his own swift
undercuts were fain to give the other an attack of indigestion. Yet each
stood firmly in his place not moving backward or forward a foot for a
good half hour, nor thinking of crying ``Enough!'' though some chance
blow seemed likely to knock one or the other off the narrow foot-bridge.
The giant's face was getting red, and his breath came snorting forth
like a bull's. He stepped forward with a furious onslaught to finish
this audacious fellow. Robin dodged his blows lightly, then sprang in
swiftly and unexpectedly and dealt the stranger such a blow upon the
short ribs that you would have sworn the tanner was trimming down his
hides for market.

The stranger reeled and came within an ace of falling, but regained his
footing right quickly.

``By my life, you can hit hard!'' he gasped forth, giving back a blow
almost while he was yet staggering.

This blow was a lucky one. It caught Robin off his guard. His stick had
rested a moment while he looked to see the giant topple into the water,
when down came the other upon his head, whack! Robin saw more stars in
that one moment than all the astronomers have since discovered, and
forthwith he dropped neatly into the stream.

The cool rushing current quickly brought him to his senses, howbeit he
was still so dazed that he groped blindly for the swaying reeds to pull
himself up on the bank. His assailant could not forbear laughing
heartily at his plight, but was also quick to lend his aid. He thrust
down his long staff to Robin crying, ``Lay hold of that, an your fists
whirl not so much as your head!''

Robin laid hold and was hauled to dry land for all the world like a
fish, except that the fish would never have come forth so wet and
dripping. He lay upon the warm bank for a space to regain his senses.
Then he sat up and gravely rubbed his pate.

``By all the saints!'' said he, ``you hit full stoutly. My head hums
like a hive of bees on a summer morning.''

Then he seized his horn, which lay near, and blew thereon three shrill
notes that echoed against the trees. A moment of silence ensued, and
then was heard the rustling of leaves and crackling of twigs like the
coming of many men; and forth from the glade burst a score or two of
stalwart yeomen, all clad in Lincoln green, like Robin, with good Will
Stutely and the widow's three sons at their head.

``Good master,'' cried Will Stutely, ``how is this? In sooth there is
not a dry thread on your body.''

``Why, marry,'' replied Robin, ``this fellow would not let me pass the
footbridge, and when I tickled him in the ribs, he must needs answer by
a pat on the head which landed me overboard.''

``Then shall he taste some of his own porridge,'' quoth Will. ``Seize
him, lads!''

``Nay, let him go free,'' said Robin. ``The fight was a fair one and I
abide by it. I surmise you also are quits?'' he continued, turning to
the stranger with a twinkling eye.

``I am content,'' said the other, ``for verily you now have the best end
of the cudgel. Wherefore, I like you well, and would fain know your
name.''

``Why,'' said Robin, ``my men and even the Sheriff of Nottingham know me
as Robin Hood, the outlaw.''

``Then am I right sorry that I beat you,'' exclaimed the man, ``for I
was on my way to seek you and to try to join your merry company. But
after my unmannerly use of the cudgel, I fear we are still strangers.''

``Nay, never say it!'' cried Robin, ``I am glad I fell in with you;
though, sooth to say, I did all the falling!''

And amid a general laugh the two men clasped hands, and in that clasp
the strong friendship of a lifetime was begun.

``But you have not yet told us your name,'' said Robin, bethinking
himself.

``Whence I came, men call me John Little.''

``Enter our company then, John Little; enter and welcome. The rites are
few, the fee is large. We ask your whole mind and body and heart even
unto death.''

``I give the bond, upon my life,'' said the tall man.

Thereupon Will Stutely, who loved a good jest, spoke up and said: ``The
infant in our household must be christened, and I'll stand godfather.
This fair little stranger is so small of bone and sinew, that his old
name is not to the purpose.'' Here he paused long enough to fill a horn
in the stream. ``Hark ye, my son,''--standing on tiptoe to splash the
water on the giant--``take your new name on entering the forest. I
christen you Little John.''

At this jest the men roared long and loud.

``Give him a bow, and find a full sheath of arrows for Little John,''
said Robin joyfully. ``Can you shoot as well as fence with the staff, my
friend?''

``I have hit an ash twig at forty yards,'' said Little John.

Thus chatting pleasantly the band turned back into the woodland and
sought their secluded dell, where the trees were the thickest, the moss
was the softest, and a secret path led to a cave, at once a retreat and
a stronghold. Here under a mighty oak they found the rest of the band,
some of whom had come in with a brace of fat does. And here they built a
ruddy fire and sat down to the meat and ale, Robin Hood in the center
with Will Stutely on the one hand and Little John on the other. And
Robin was right well pleased with the day's adventure, even though he
had got a drubbing; for sore ribs and heads will heal, and 'tis not
every day that one can find a recruit as stout of bone and true of soul
as Little John.
