\chapter{How Robin Hood Was Tanned of the Tanner}

\begin{quote}
In Nottingham there lived a jolly tanner,\\
With a hey down, down, a down down!\\
His name was Arthur-a-Bland,\\
There was ne'er a squire in Nottinghamshire

Dare bid bold Arthur stand.\\
And as he went forth, in a summer's morning,\\
With a hey down, down, a down down!\\
To the forest of merrie Sherwood,\\
To view the red deer, that range here and there,\\
There met he with bold Robin Hood.
\end{quote}

\lettrine{T}{he} Sheriff's daughter bided for several days in the faint
hope that she might hear tidings of the prattling tinker. But never a
word heard she, and she was forced to the conclusion that her messenger
had not so much as laid eyes upon the outlaw. Little recked she that he
was, even then, grinding sword-points and sharpening arrows out in the
good greenwood, while whistling blithely or chatting merrily with the
good Friar Tuck.

Then she bethought herself of another good man, one Arthur-a-Bland, a
tanner who dwelt in Nottingham town and was far-famed in the tourneys
round about. He had done some pretty tricks at archery, but was
strongest at wrestling and the quarter-staff. For three years he had
cast all comers to the earth in wrestling until the famous Eric o'
Lincoln broke a rib for him in a mighty tussle. Howsoever, at
quarter-staff he had never yet met his match; so that there was never a
squire in Nottinghamshire dare bid bold Arthur stand.

\begin{quote}
With a long pike-staff on his shoulder,\\
So well he could clear his way\\
That by two and three he made men flee\\
And none of them could stay.
\end{quote}

Thus at least runs the old song which tells of his might.

``This is just the man for me!'' thought the Sheriff's daughter to
herself; and she forthwith summoned him to the Mansion House and
commissioned him to seek out Robin Hood.

The warrant was quite to Arthur's liking, for he was happiest when out
in the forest taking a sly peep at the King's deer; and now he reckoned
that he could look at them boldly, instead of by the rays of the moon.
He could say to any King's Forester who made bold to stop him: ``I am
here on the King's business!''

``Gramercy! No more oak-bark and ditch-water and the smell of
half-tanned hides to-day!'' quoth he, gaily. ``I shall e'en see what the
free air of heaven tastes like, when it sweeps through the open wood.''

So the tanner departed joyfully upon his errand, but much more
interested in the dun deer of the forest than in any two-legged rovers
therein. This interest had, in fact, caused the Foresters to keep a
shrewd eye upon him in the past, for his tannery was apt to have plenty
of meat in it that was more like venison than the law allowed. As for
the outlaws, Arthur bore them no ill-will; indeed he had felt a secret
envy in his heart at their free life; but he was not afraid to meet any
two men who might come against him. Nathless, the Sheriff's daughter did
not choose a very good messenger, as you shall presently see.

Away sped the tanner, a piece of bread and some wine in his wallet, a
good longbow and arrows slung across his shoulder, his stout
quarter-staff in his hand, and on his head a cap of trebled raw-hide so
tough that it would turn the edge of a broadsword. He lost no time in
getting out of the hot sun and into the welcome shade of the forest,
where he stalked cautiously about seeking some sign of the dun deer.

Now it so chanced that upon that very morning Robin Hood had sent Little
John to a neighboring village to buy some cloth of Lincoln green for new
suits for all the band. Some of the money recently won of the King was
being spent in this fashion, `gainst the approach of winter. Will
Scarlet had been sent on a similar errand to Barnesdale some time
before, if you remember, only to be chased up the hill without his
purchase. So to-day Little John was chosen, and for sweet company's sake
Robin went with him a part of the way until they came to the ``Seven
Does,'' the inn where Robin had recently played his prank upon Middle
the tinker. Here they drank a glass of ale to refresh themselves withal,
and for good luck; and Robin tarried a bit while Little John went on his
errand.

Presently Robin entered the edge of the wood, when whom should he see
but Arthur-a-Bland, busily creeping after a graceful deer that browsed
alone down the glade. ``Now by Saint George and the Dragon!'' quoth
Robin to himself. ``I much fear that yon same fellow is a rascally
poacher come after our own and the King's meat!''

For you must know, by a curious process of reasoning, Robin and his men
had hunted in the royal preserves so long that they had come to consider
themselves joint owners to every animal which roamed therein.

``Nay!'' he added, ``this must be looked into! That cow-skin cap in
sooth must hide a scurvy varlet!''

And forthwith he crept behind a tree, and thence to another, stalking
our friend Arthur as busily as Arthur was stalking the deer.

This went on for quite a space, until the tanner began to come upon the
deer and to draw his bow in order to tickle the victim's ribs with a
cloth-yard shaft. But just at this moment Robin unluckily trod upon a
twig which snapped and caused the tanner to turn suddenly.

Robin saw that he was discovered, so he determined to put a bold face on
the matter, and went forward with some smart show of authority.

``Hold!'' he cried: ``stay your hand! Why, who are you, bold fellow, to
range so boldly here? In sooth, to be brief, ye look like a thief that
has come to steal the King's deer.''

``Marry, it is scant concern of yours, what I look like!'' retorted
Arthur-a-Bland. ``Who are you, who speak so bravely?''

``You shall soon find out who I am!'' quoth Robin, determining to find
some sport in the matter. ``I am a keeper of this forest. The King knows
that I am looking after his deer for him; and therefore we must stay
you.''

``Have you any assistants, friend?'' asked the tanner calmly. ``For it
is not one man alone who can stop me.''

``Nay truly, gossip,'' replied Robin. ``I have a good yew bow, also a
right sharp blade at my side. Nathless I need no better assistant than a
good oak-graff like unto yours. Give me a baker's dozen of minutes with
it and it shall pleasure me to crack that pate of yours for your
sauciness!''

``Softly, my man! Fair and softly! Big words never killed so much as a
mouse--least of all yon deer which has got away while you were filling
all the woods with your noisy breath. So choose your own playthings. For
your sword and your bow I care not a straw; nor for all your arrows to
boot. If I get but a knock at you, `twill be as much as you'll need.''

``Now by our Lady! Will you listen to the braggart?'' cried Robin in a
fine rage. ``Marry, but I'll teach ye to be more mannerly!''

So saying he unbuckled his belt; and, flinging his bow upon the ground
he seized hold of a young sapling that was growing near by. His hunting
knife soon had it severed and lopped into shape.

``Now come, fellow!'' said Arthur-a-Bland, seeing that he was ready.
``And if I do not tan your hide for you in better shape than ever
calf-skin was turned into top-boots, may a murrain seize me!''

``Stay,'' said Robin, ``methinks my cudgel is half a foot longer than
yours. I would have them of even length before you begin your tanning.''

``I pass not for length,'' bold Arthur replied; ``my staff is long
enough, as you will shortly find out. Eight foot and a half, and `twill
knock down a calf''--here he made it whistle in the air--``and I hope it
will knock down you.''

Forthwith the two men spat on their hands, laid firm hold upon their
cudgels and began slowly circling round each other, looking for an
opening.

Now it so chanced that Little John had fared expeditiously with his
errand. He had met the merchant, from whom he was wont to buy Lincoln
green, coming along the road; and had made known his wants in few words.
The merchant readily undertook to deliver the suits by a certain day in
the following month. So Little John, glad to get back to the cool
shelter of the greenwood, hasted along the road lately taken by Robin.

Presently he heard the sound of angry voices, one of which he recognized
as his captain's.

``Now, Heaven forfend,'' quoth he, ``that Robin Hood has fallen into the
clutches of a King's man! I must take a peep at this fray.''

So he cautiously made his way from tree to tree, as Robin had done, till
he came to the little open space where Robin and Arthur were circling
about each other with angry looks, like two dogs at bay.

``Ha! this looks interesting!'' muttered Little John to himself, for he
loved a good quarter-staff bout above anything else in the world, and
was the best man at it in all the greenwood. And he crawled quietly
underneath a friendly bush--much as he had done when Robin undertook to
teach Will Scarlet a lesson--and chuckled softly to himself and slapped
his thigh and prepared to watch the fight at his ease.

Indeed it was both exciting and laughable. You would have chuckled one
moment and caught your breath the next, to see those two stout fellows
swinging their sticks--each half as long again as the men were, and
thick as their arm--and edging along sidewise, neither wishing to strike
the first blow.

At last Robin could no longer forbear, and his good right arm swung
round like a flash. Ping! went the stick on the back of the other's
head, raising such a welt that the blood came. But the tanner did not
seem to mind it at all, for bing! went his own staff in return, giving
Robin as good as he had sent. Then the battle was on, and furiously it
waged. Fast fell the blows, but few save the first ones landed, being
met in mid-air by a counter-blow till the thwacking sticks sounded like
the steady roll of a kettle-drum and the oak--bark flew as fine as it
had ever done in Arthur-a-Bland's tannery.

Round and round they fought, digging their heels into the ground to keep
from slipping, so that you would have vowed there had been a yoke of
oxen ploughing a potato-patch. Round and round, up and down, in and out,
their arms working like threshing-machines, went the yeoman and the
tanner, for a full hour, each becoming more astonished every minute that
the other was such a good fellow. While Little John from underneath his
bushy covert had much ado to keep from roaring aloud in pure joy.

Finally Robin saw his chance and brought a full arm blow straight down
upon the other's head with a force that would have felled a bullock. But
Arthur's trebled cow-skin cap here stood him in good stead: the blow
glanced off without doing more than stunning him. Nathless, he reeled
and had much ado to keep from falling; seeing which Robin stayed his
hand--to his own sorrow, for the tanner recovered his wits in a
marvelous quick space and sent back a sidelong blow which fairly lifted
Robin off his feet and sent him tumbling on to the grass.

``Hold your hand! hold your hand!'' roared Robin with what little breath
he had left. ``Hold, I say, and I will give you the freedom of the
greenwood.''

``Why, God-a-mercy,'' said Arthur; ``I may thank my staff for that--not
\emph{you}.''

``Well, well, gossip' let be as it may. But prithee tell me your name
and trade. I like to know fellows who can hit a blow like that same
last.''

``I am a tanner,'' replied Arthur-a-Bland. ``In Nottingham long have I
wrought. And if you'll come to me I swear I'll tan your hides for
naught.''

``Odds bodikins!'' quoth Robin ruefully. ``Mine own hide is tanned
enough for the present. Howsoever, there be others in this wood I would
fain see you tackle. Harkee, if you will leave your tan-pots and come
with me, as sure as my name is Robin Hood, you shan't want gold or
fee.''

``By the breath o' my body!'' said Arthur, ``that will I do!'' and he
gripped him gladly by the hand. ``But I am minded that I clean forgot
the errand that brought me to Sherwood. I was commissioned by some,
under the Sheriff's roof, to capture you.''

``So was a certain tinker, now in our service,'' said Robin smilingly.

``Verily 'tis a new way to recruit forces!'' said the tanner laughing
loudly. ``But tell me, good Robin Hood, where is Little John? I fain
would see him, for he is a kinsman on my mother's side.''

``Here am I, good Arthur-a-Bland!'' said a voice; and Little John
literally rolled out from under the bush to the sward. His eyes were
full of tears from much laughter which had well-nigh left him powerless
to get on his feet.

As soon as the astonished tanner saw who it was, he gave Little John a
mighty hug around the neck, and lifted him up on his feet, and the two
pounded each other on the back soundly, so glad were they to meet again.

``O, man, man!'' said Little John as soon as he had got his breath.
``Never saw I so fine a sight in all my born days. You did knock him
over like as he were a ninepin!''

``And you do joy to see me thwacked about on the ribs?'' asked Robin
with some choler.

``Nay, not that, master!'' said Little John. ``But 'tis the second time
I have had special tickets to a show from beneath the bushes, and I
cannot forbear my delight. Howsoever, take no shame unto yourself, for
this same Arthur-a-Bland is the best man at the quarter-staff in all
Nottinghamshire. It commonly takes two or three men to hold him.''

``Unless it be Eric o' Lincoln,'' said Arthur modestly; ``and I well
know how you paid him out at the Fair.''

``Say no more!'' said Robin springing to his feet; ``for well I know
that I have done good business this day, and a few bruises are easy
payment for the stout cudgel I am getting into the band. Your hand
again, good Arthur-a-Bland! Come! let us after the deer of which I
spoiled your stalking.''

``Righty gladly!'' quoth Arthur. ``Come, Cousin Little John! Away with
vats and tan-bark and vile-smelling cowhides! I'll follow you two in the
sweet open air to the very ends of earth!''
